---
title: "Turtles tidied data"
author: "Eric Coyle"
date: "2/21/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(shiny)
library(paletteer)
library(ggmap)
library(rgeos)
library(rgdal)
library(devtools)
```


```{r}
turtles<-read.csv(here("Final Project","Official Final Project Documents/Turtle_data.csv"))
```

```{r}
turtles_tidy<-turtles%>%
  na_if("")%>%
  na_if("0")%>%
  na_if("0.0")%>%
  mutate(dead_alive_new=if_else(Dead_Alive=="alive","Alive",Dead_Alive))
```

```{r}
library(lubridate)
```

```{r}
turtles3<-turtles_tidy%>%
  mutate(DateCapture_new=dmy(DateCapture))%>%
  filter(!is.na(DateCapture_new))%>%
  mutate(Capture_month=month(DateCapture_new))%>%
  mutate(Capture_day=day(DateCapture_new))%>%
  mutate(capture_week_day=wday(DateCapture_new))
turtles3<-clean_names(turtles3)
head(turtles3)
view(turtles3)
```

#mapping for release and capture sites

```{r}
turtles3%>%
  select(cap_latitude,cap_longitude)%>%
  summary()
```

```{r}
cap_lat <- c(33.49, 38.41)
cap_long <- c(-79.08, -75.47)
bbox <- make_bbox(cap_long, cap_lat, f = 0.1)
```

```{r}
cap_map_base <- get_map(bbox, maptype = "terrain-background", source = "stamen")
ggmap(cap_map_base)
```



```{r}
ggmap(cap_map_base) + 
  geom_point(data = turtles3, aes(cap_longitude,cap_latitude,color=species,shape=dead_alive_new), size = 2, alpha = 0.7) +
           labs(x = "Longitude", y = "Latitude", title = "Capture Locations")
```
```{r}
turtles3%>%
  select(rel_latitude,rel_longitude)%>%
  summary()
```

```{r}
rel_lat <- c(27.82, 39)
rel_long <- c(-81.24, -36.03)
bbox2 <- make_bbox(rel_long, rel_lat, f = 0.05)
```

```{r}
rel_map_base <- get_map(bbox2, maptype = "terrain-background", source = "stamen")
ggmap(rel_map_base)
```
```{r}
ggmap(rel_map_base) + 
  geom_point(data = turtles3, aes(rel_longitude,rel_latitude,color=species), size = 2, alpha = 0.7) +
           labs(x = "Longitude", y = "Latitude", title = "Release Locations")
```
#we can then make additional maps with color or shape of the dots for different variables. We also should pick a unifying color scheme/text/chart and map theme for all of our images (Ideally a color that stands out against the terrain background if we want to stick with terrain for our maps)

#The following are some ideas for graphs that I thought would be good to include, but we can each come up and explore some of our own to compile into the final product

```{r}
turtles3%>%
  group_by(year)%>%
  summarise(n())
```
```{r}
turtles3%>%
  group_by(year)%>%
  ggplot(aes(x=year,fill=dead_alive_new))+
  geom_bar(position = "dodge")
```


```{r}
turtles3%>%
  group_by(capture_month)%>%
  summarise(n())
```

```{r}
turtles3%>%
  ggplot(aes(month(x=date_capture_new,label=TRUE)))+
  geom_bar(position = "dodge")+
  labs(x = NULL,
         y = "Number of Captures")
```


```{r}
turtles3%>%
  group_by(capture_day)%>%
  summarise(n())
```

```{r}
turtles3%>%
  ggplot(aes(day(x=date_capture_new)))+
  geom_density(fill="green")+
  labs(x = NULL,
         y = "Number of Captures")
```


```{r}
turtles3%>%
  group_by(capture_week_day)%>%
  summarise(n())
```
```{r}
turtles3%>%
  ggplot(aes(wday(x=date_capture_new,label=TRUE)))+
  geom_bar()+
  labs(x = NULL,
         y = "Number of Captures")
```
#Monday and Friday are popular trapping days apparently

```{r}
turtles3%>%
  group_by(year)%>%
  mutate(mean_body_length=mean(scl_tip,na.rm=T))%>%
  ggplot(aes(x=year,y=mean_body_length,group=1,color=species))+
  geom_line()+
  geom_point()+
  facet_wrap(~species)
```
```{r}
turtles3%>%
  ggplot(aes(x=species,y=ccl_tip,color=species))+
  geom_boxplot()
```
#we of course can fix the axis and replicate this kind of graph for other measurements (SCL_noth/SCl_tip, CCl_notch, etc)

```{r}
turtles3%>%
  filter(gear!=""&species!="")%>%
  ggplot(aes(x=gear,fill=species))+
  geom_bar(position = "dodge")+
  labs(title = "Gear used for Different Species",x="Gear",y="Number Captured with Given Method")+
  theme(axis.text.x = element_text(angle = 60,hjust = 1))+
  scale_y_log10()
```
```{r}
names(turtles3)
```
#This App is for some categorical variables across the years for different spcecies
```{r}
library(shinydashboard)
```


```{r}
turtles3$year<-as.character(turtles3$year)
ui <- dashboardPage(skin="green",
  dashboardHeader(title = "Turtle Captures"),
  dashboardSidebar(disable = F),
  dashboardBody(selectInput("species", "Select Species:", 
                  choices=unique(turtles3$species)),
  fluidRow(
  box(title = "Plot Options", width = 4,
  selectInput("x", "Catch Details", choices = c("research_type","dead_alive_new", "body_area_pit","body_area","record_type","cap_region","rel_region"), 
              selected = "record_type"),
      hr(),
      helpText("Source: (https://www.fisheries.noaa.gov/inport/item/35875). Capture efforts were conducted to evaluate the growth rates, sex ratios, size distribution, species composition, genetic composition, relative survival rates and foraging ecology of sea turtle populations in NC.")
  ), 
  box(title = "Turtle Information", width = 8,
  plotOutput("plot", width = "800px", height = "500px")
  ) 
  ) 
  ) 
  )
server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
  turtles3 %>%
      filter(species == input$species) %>%
  ggplot(aes_string(x ="year",fill = input$x)) +
  geom_bar(position = "dodge")+
       scale_fill_brewer(palette = "Set1")+
  theme_light(base_size = 18)+
     theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(title = "Turtle Catch Characterisitics",x=NULL,y="Number of Turtles")
  })
  
  session$onSessionEnded(stopApp)
  }

shinyApp(ui, server)
```

```{r}
turtles3%>%
  tabyl(haplotype)
```
#This next app shows which tests were perfomed across the years.

```{r}
turtles3$year<-as.character(turtles3$year)
ui <- dashboardPage(skin="green",
  dashboardHeader(title = "Turtle Capture Tests"),
  dashboardSidebar(disable = F),
  dashboardBody(selectInput("species", "Select Species:", 
                  choices=unique(turtles3$species)),
  fluidRow(
  box(title = "Plot Options", width = 4,
  selectInput("x", "Type of Test", choices = c("oxtetracyclene","health_blood","satellite_tag","holding_facility","sex_laparoscopy",     "sex_necropsy","sex_testosterone_level_1","sex_testosterone_level_2","hematology","oc_ob","fh","metals","sia_skin","sia_bone",     "sia_blood","sia_barnacles","sia_scutes","sia_analyzed","photos","scute","cloacal","lesion","fat","fecal","pathogens",     "entangled","imaging","organ_biopsy"), 
              selected = "entangled"),
      hr(),
      helpText("Source: (https://www.fisheries.noaa.gov/inport/item/35875). Capture efforts were conducted to evaluate the growth rates, sex ratios, size distribution, species composition, genetic composition, relative survival rates and foraging ecology of sea turtle populations in NC.")
  ), 
  box(title = "Test Performed T/F", width = 8,
  plotOutput("plot", width = "800px", height = "500px")
  ) 
  ) 
  ) 
  )
server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
  turtles3 %>%
      filter(species == input$species) %>%
  ggplot(aes_string(x ="year",fill = input$x)) +
  geom_bar(position = "dodge")+
       scale_fill_brewer(palette = "Set1")+
  theme_light(base_size = 18)+
     theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(title = "Tests Performed On Turtles",x=NULL,y="Number of Turtles Tested for Selected Test")
  })
  
  session$onSessionEnded(stopApp)
  }

shinyApp(ui, server)
```

Quantitative Data Apps (rought first idea, will adjust aes on Thursday)
```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Turtle Measurements"),
  dashboardSidebar(disable = T),
  dashboardBody(
  fluidRow(
  box(
  selectInput("x", "Select X Variable", choices = c("scl_notch ", "scl_tip", "scw", "ccl_notch", "ccl_tip", "ccw", "circumference", "tail", "girth", "depth_mid", "weight"), selected = "scl_notch"),
  selectInput("y", "Select Y Variable", choices = c("scl_notch ", "scl_tip", "scw", "ccl_notch", "ccl_tip", "ccw", "circumference", "tail", "girth", "depth_mid", "weight"), selected = "scl_notch"),
  ), # close the first box
  box(
  plotOutput("plot", width = "500px", height = "500px")
  ) 
  ) 
  ) 
) 

server <- function(input, output, session) { 
  output$plot <- renderPlot({
  ggplot(turtles3, aes_string(x = input$x, y = input$y)) + geom_point(alpha=0.8) + theme_light(base_size = 18)
  })
  session$onSessionEnded(stopApp)
  }

shinyApp(ui, server)
```
```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Turtle Measurements"),
  dashboardSidebar(disable = T),
  dashboardBody(
  fluidRow(
  box(
  selectInput("x", "Select X Variable", choices = c("cap_latitude", "cap_longitude"), selected = "scl_notch"),
  selectInput("y", "Select Y Variable", choices = c("scl_notch ", "scl_tip", "scw", "ccl_notch", "ccl_tip", "ccw", "circumference", "tail", "girth", "depth_mid", "weight"), selected = "scl_notch"),
  ), # close the first box
  box(
  plotOutput("plot", width = "500px", height = "500px")
  ) 
  ) 
  ) 
) 

server <- function(input, output, session) { 
  output$plot <- renderPlot({
  ggplot(turtles3, aes_string(x = input$x, y = input$y)) + geom_point(alpha=0.8) + theme_light(base_size = 18)
  })
  session$onSessionEnded(stopApp)
  }

shinyApp(ui, server)
```
#My second app was an interesting idea, but it doesn't come out that pretty. Might just make some new maps where I scale point size by turtle weight, size, or depth. 

Additional Graph Ideas: 
```{r}
turtles3 %>% 
    mutate(mass_category = case_when(weight <= 30 ~ "small", weight >= 30 ~ "large")) %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, fill = mass_category))+
  geom_bar()+
  labs(title = "Turtle Capture by Weight Class",
       x = "Year",
       y= "n")+
    theme_linedraw()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1), plot.title = element_text(size = rel(1.5), hjust = 0.5))
```
#test data

